---
title: "M4Comp Benchmark Methods Part 2: Automate Benchmark Applications"
author: "Timo Meiendresch"
date: "06/06/2019"
knit: (function(input_file, encoding) {
  out_dir <- 'html_outputs';
  rmarkdown::render(input_file,
  encoding=encoding,
  output_file=file.path(dirname(input_file), out_dir, 'M4Comp_Benchmarks2.html'))})
---

# Introduction

In a next step we'll use the previously introduced methods saved in the script `my_utils.R` and apply them to more than one series. The objective is to apply the benchmark methods to a huge dataset (i.e. 100,000 series) and automation using functional programming is required. 

The progression is to first apply the benchmark methods to a smaller subset of 3 time series with different frequencies. Afterwards, we'll use a bigger subset of Monthly (Yearly, Quarterly) data.

```{r, message=FALSE, echo=FALSE}
rm(list=ls())
graphics.off()
```

```{r, message=FALSE}
# standard libraries
library(forecast)
library(ggplot2)
library(M4comp2018)
```

## Load functions from Script




