---
title: "M4Comp Benchmark Methods Part 2: Automate Benchmark Methods"
author: "Timo Meiendresch"
date: "06/06/2019"
knit: (function(input_file, encoding) {
  out_dir <- 'html_outputs';
  rmarkdown::render(input_file,
  encoding=encoding,
  output_file=file.path(dirname(input_file), out_dir, 'M4Comp_Benchmarks2.html'))})
---

# Introduction

In a next step we'll use the previously introduced methods saved in the script `my_utils.R` and apply them to more than one series. The objective is to apply the benchmark methods to a huge dataset (i.e. 100,000 series) and automation using abstraction is necessary. 

The progression is to first apply the benchmark methods to a smaller subset of 3 time series with different frequencies. Afterwards, we'll use a bigger subset of Monthly (Yearly, Quarterly, ...) data.

```{r, message=FALSE, echo=FALSE}
rm(list=ls())
graphics.off()
```

```{r, message=FALSE}
# standard libraries
library(forecast)
library(ggplot2)
library(M4comp2018)
```

# Introduction

In the following document we'll apply the benchmark methods on the M4Comp data. We'll start with a small subsample of data that is randomly drawn from the monthly time series. 

## Random sample 

```{r}
# load data
data(M4)

# extract monthly series 
M4monthly <- Filter(function(l) l$period=="Monthly", M4)

# set a random seed
set.seed(16)

# determine sample size
sample_n <- 10
idx <- sample(1:length(M4monthly), size = 10)

# draw sample from monthly data
M4monthly_sub <- M4monthly[idx]

# clear-memory-except
rm(list=setdiff(ls(), ls()[grep(pattern=c("M4monthly_sub"), x=ls())] ) )
```

### Loading functions

The required functions are saved in `my_utils.R` and will now be loaded for usage. 

```{r}
# load functions
source("my_utils.R")
```

### Create a container for the sMAPE and MASE data

```{r}
# benchmark names 
Names_benchmarks <- c("Naive", "sNaive", "Naive2", "SES", "Holt", "Damped", "Theta", "Comb")

# train-test split 
data_train <- data_test <- M4monthly_sub
fh <- M4monthly_sub[[1]]$h
n_fc <- length(Names_benchmarks)
n_series <- length(data_train)

# Container for sMAPE, MASE
Total_MASE <- array(NA,dim = c(n_fc, fh, n_series))
Total_sMAPE <- Total_MASE
```

```{r}
# Example of fourth serise in the subset
nr <- 4
insample <- data_train[[nr]]$x
outsample <- data_test[[nr]]$xx
fc <- benchmarks(input = insample, fh=fh)

# Seasonality
SeasonalityTest(insample, fh)

autoplot(window(insample, start=2012)) +
  autolayer(outsample, series="test") + 
  autolayer(fc$Naive, series="Naive") +
  autolayer(fc$sNaive, series="sNaive") +
  autolayer(fc$Holt, series="Holt") +
  ggtitle("Example plot of series #4") +
  xlab("Year") + ylab("value") +
  guides(colour=guide_legend(title="Forecast"))
```

Note that if `SeasonalityTest` evaluates to `TRUE`, then the following methods (including their forecasts) are adjusted using the seasonal component: 

- `Naive`
- `SES` 
- `Holt`
- `Damped`: Also known as **Holt method with damped trend**

## Results 

```{r}
for (i in 1:length(data_train)){
  insample <- data_train[[i]]$x
  outsample <- data_test[[i]]$xx
  # forecasts
  fc <- benchmarks(input = insample, fh = fh)
  
  # sMAPE
  for(j in 1:n_fc){
    Total_sMAPE[j,,i] <- cal_sMAPE(outsample, fc[[j]])
  }
  
  # MASE
  for (j in 1:n_fc){
    Total_MASE[j,,i] <- cal_MASE(insample, outsample, fc[[j]])
  }
}
```

### Results sMAPE

```{r}
# sMAPE
for (i in 1:n_fc){
  print(paste(Names_benchmarks[i],"-", round(mean(Total_sMAPE[i,,]),4)))
}
```

### Results MASE

```{r}
# MASE 
for (i in 1:n_fc){
  print(paste(Names_benchmarks[i], "-", round(mean(Total_MASE[i,,]), 4)))
}
```

### Results OWA

```{r}
# OWA 
for(i in 1:n_fc){
  rel_MASE <- (mean(Total_MASE[i,,]) / mean(Total_MASE[3,,]))
  rel_sMAPE <- (mean(Total_sMAPE[i,,])/ mean(Total_sMAPE[3,,]))
  owa <- (rel_MASE + rel_sMAPE)/2
  print(paste(Names_benchmarks[i], "-",round(owa,4)))
}
```



